---
title: "Mime_analysis"
author: "Poojitha Kolli, Dillen Wischmeier"
date: "2025-05-03"
output: html_document
---

# Set the working directory

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/kpooj/Documents/MS/Fourth_Sem/ML assignments/Project/git-repo/")
```

# Load the data

```{r}
library(dplyr)

# Survival data
survival_data <- read.delim("./results/tcga_lusc_survival_data.tsv", stringsAsFactors = FALSE)

# Filtered data
tcga_filtered <- readRDS("./results/tcga_filtered_expression_normalized.rds")

# Full expression data
tcga_full <- readRDS("./results/tcga_full_expression_normalized.rds")
```

# Mime Input

### Filtered expression data

```{r}
# Filtered expression data
# Full expression data
lusc_filtered <- tcga_filtered %>%
  as.data.frame() %>%
  select(1:468) %>%   # lusc samples
  rename_with(~ substr(., 1, 12)) %>%
  t() %>%             # transpose
  as.data.frame()

# Reorder lusc_full to match the samples in survival_data
lusc_filtered <- lusc_filtered[match(survival_data$case_submitter_id, rownames(lusc_filtered)), , drop = FALSE]

# Combine survival data with expression data
lusc_filtered <- cbind(survival_data, lusc_filtered)
colnames(lusc_filtered)[colnames(lusc_filtered) == "case_submitter_id"] <- "ID"

# sampling
set.seed(2025) # for reproducibility
train_indices <- sample(1:nrow(lusc_filtered), round(0.75 * nrow(lusc_filtered)))

# Create training and testing datasets
lusc_train <- lusc_filtered[train_indices, ]
lusc_test <- lusc_filtered[-train_indices, ]

# Save both datasets to an .RData file
split_data <- list(train = lusc_train, test = lusc_test)
dir.create("./Mime/Input", recursive = TRUE, showWarnings = FALSE)
save(split_data, file = "./Mime/Input/lusc_split_filtered.RData")
```

### Full expression data

```{r}
# Full expression data
lusc_full <- tcga_full %>%
  as.data.frame() %>%
  select(1:468) %>%   # lusc samples
  rename_with(~ substr(., 1, 12)) %>%
  t() %>%             # transpose
  as.data.frame()

# Reorder lusc_full to match the samples in survival_data
lusc_full <- lusc_full[match(survival_data$case_submitter_id, rownames(lusc_full)), , drop = FALSE]

# Combine survival data with expression data
lusc_full <- cbind(survival_data, lusc_full)
colnames(lusc_full)[colnames(lusc_full) == "case_submitter_id"] <- "ID"

# sampling
set.seed(2025) # for reproducibility
train_indices <- sample(1:nrow(lusc_full), round(0.75 * nrow(lusc_full)))

# Create training and testing datasets
lusc_train_full <- lusc_full[train_indices, ]
lusc_test_full <- lusc_full[-train_indices, ]

# Save both datasets to an .RData file
split_data_lusc <- list(train = lusc_train_full, test = lusc_test_full)
save(split_data_lusc, file = "./Mime/Input/lusc_split_full.RData")
```

```{r}
# --- Cleanup workspace ---
rm(list = ls())
gc()  # garbage collection to free memory
```


# Mime_analysis
# Using the highly variable genes dataset (~27k genes)-------------------------

## Load the data

```{r}
# Load required libraries
library(Mime1)
library(survminer)
library(immunedeconv)

# Load necessary data
load("./Mime/Input/lusc_split_filtered.RData")
gene_list <- readRDS("./results/kegg_pi3k_common_genelist.rds")
split_data[["train"]]$OS.time[split_data[["train"]]$OS.time == 0] <- 1e-5
```


## 1. Select the Optimal Model

```{r}
# Perform model selection
res <- ML.Dev.Prog.Sig(
  train_data = split_data$train,
  list_train_vali_Data = split_data,
  unicox.filter.for.candi = TRUE,
  unicox_p_cutoff = 0.05,
  candidate_genes = gene_list,
  mode = 'all',
  nodesize = 5,
  seed = 5201314
)

# Check warnings after model fitting
warnings()
```

### Plot C-index of Each Model

```{r}
# Plot the C-index of each model
plot_cindex_all = cindex_dis_all(res, validate_set = names(split_data)[-1], order = names(split_data), width = 0.35)
dir.create("./Mime/Ouput_27k", recursive = TRUE, showWarnings = FALSE)
ggsave("./Mime/Ouput_27k/cindex_all_models.png", plot = plot_cindex_all, width = 6.25, height = 18, dpi = 400)
```

### Plot C-index for a Specific Model

```{r}
# Plot the C-index for a specific model among different datasets
cindex_select_plot = cindex_dis_select(res,
                                       model = "Lasso + SuperPC",
                                       order = names(split_data))
ggsave("./Mime/Ouput_27k/cindex_selected_model.png", plot = cindex_select_plot, width = 6, height = 1.5, dpi = 300)
```

```{r}
# --- Cleanup workspace ---
rm(list = ls())
gc()
```



# Using whole dataset of cancer samples (~60k genes) ---------------------------

## Load the data

```{r}
# Load necessary data
load("./Mime/Input/lusc_split_full.Rdata")
gene_list <- readRDS("./results/kegg_pi3k_common_genelist.rds")
split_data_lusc[["train"]]$OS.time[split_data_lusc[["train"]]$OS.time == 0] <- 1e-5
```

## 1. Select the Optimal Model

```{r}
# Perform model selection
res <- ML.Dev.Prog.Sig(
  train_data = split_data_lusc$train,
  list_train_vali_Data = split_data_lusc,
  unicox.filter.for.candi = TRUE,
  unicox_p_cutoff = 0.05,
  candidate_genes = gene_list,
  mode = 'all',
  nodesize = 5,
  seed = 5201314
)

# Check warnings after model fitting
warnings()
```

### Plot C-index of Each Model

```{r}
# Plot the C-index of each model
plot_cindex_all = cindex_dis_all(res, validate_set = names(split_data_lusc)[-1], order = names(split_data_lusc), width = 0.35)
dir.create("./Mime/Ouput_60k", recursive = TRUE, showWarnings = FALSE)
ggsave("./Mime/Ouput_60k/cindex_all_models.png", plot = plot_cindex_all, width = 6.25, height = 18, dpi = 400)
```

### Plot C-index for a Specific Model

```{r}
# Plot the C-index for a specific model among different datasets
cindex_select_plot = cindex_dis_select(res,
                                       model = "Lasso + SuperPC",
                                       order = names(split_data_lusc))
ggsave("./Mime/Ouput_60k/cindex_selected_model.png", plot = cindex_select_plot, width = 6, height = 1.5, dpi = 300)
```



The following steps outline how we would proceed with evaluating the model’s predictive performance and conducting immune infiltration analysis. However, since we were unable to develop a reliable model during our analysis, we have decided to halt further steps at this stage.


### Plot Survival Curves Based on Risk Score

```{r}
# Generate survival curves based on risk scores from a specific model
survplot <- vector("list", 2) 
for (i in c(1:2)) {
  survplot[[i]] <- rs_sur(res, model_name = "Lasso + SuperPC", 
                          dataset = names(split_data)[i],
                          median.line = "hv", 
                          cutoff = 0.5, 
                          conf.int = TRUE,
                          xlab = "Day", 
                          pval.coord = c(1000, 0.9))
}
png("./Mime/Ouput_27k/survival_curves.png", width = 12, height = 6, units = "in", res = 300)

# Plot survival curves in a grid
aplot::plot_list(gglist = survplot, ncol = 2)
```

## 2. Calculate AUC Scores for Different Models

```{r}
# Calculate AUC for 1-year, 3-year, and 5-year predictions
all.auc.1y <- cal_AUC_ml_res(res.by.ML.Dev.Prog.Sig = res, 
                              train_data = split_data[["train"]],
                              inputmatrix.list = split_data, 
                              mode = 'all', 
                              AUC_time = 1, 
                              auc_cal_method = "KM")
all.auc.3y <- cal_AUC_ml_res(res.by.ML.Dev.Prog.Sig = res, 
                              train_data = split_data[["train"]],
                              inputmatrix.list = split_data, 
                              mode = 'all', 
                              AUC_time = 3, 
                              auc_cal_method = "KM")
all.auc.5y <- cal_AUC_ml_res(res.by.ML.Dev.Prog.Sig = res, 
                              train_data = split_data[["train"]],
                              inputmatrix.list = split_data, 
                              mode = 'all', 
                              AUC_time = 5, 
                              auc_cal_method = "KM")

```

### Plot AUC Scores for 1-Year Prediction

```{r}
# Plot AUC scores for 1-year prediction
auc_plot_1y = auc_dis_all(all.auc.1y,
                          dataset = names(split_data),
                          validate_set = names(split_data)[-1],
                          order = names(split_data),
                          width = 0.35,
                          year = 1)
ggsave("./Mime/Ouput_27kMime_Output/auc_plot_all_1yr.png", plot = auc_plot_1y, width = 6.25, height = 18, dpi = 400)

```

### Plot ROC Curve for Specific Model

```{r}
# Plot ROC curve for a specific model across different datasets
auc_1y_select_model = roc_vis(all.auc.1y,
                              model_name = "Lasso + SuperPC",
                              dataset = names(split_data),
                              order = names(split_data),
                              anno_position = c(0.65, 0.55),
                              year = 1)
ggsave("./Mime/Ouput_27k/auc_1y_select_model.png", plot = auc_1y_select_model, width = 6, height = 6, dpi = 300)
```

### Plot AUC for 1, 3, and 5 Years

```{r}
# Plot AUC for 1, 3, and 5 years
auc_1_3_5y_model = auc_dis_select(list(all.auc.1y, all.auc.3y, all.auc.5y),
                            model_name = "Lasso + SuperPC",
                            dataset = names(split_data),
                            order = names(split_data),
                            year = c(1, 3, 5))
ggsave("./Mime/Ouput_27k/auc_1_3_5y_selected_model.png", plot = auc_1_3_5y_model, width = 6, height = 1.5, dpi = 300)
```

## 3. Meta-Analysis of Univariate Cox Regression

```{r}
library(conflicted)
# Perform univariate Cox regression for a specific model
unicox.rs.res <- cal_unicox_ml_res(res.by.ML.Dev.Prog.Sig = res,
                                   optimal.model = "Lasso + SuperPC",
                                   type = 'categorical')

# Perform meta-analysis on the univariate Cox results
metamodel <- cal_unicox_meta_ml_res(input = unicox.rs.res)

# Visualize the results of the meta-analysis
meta_unicox_vis(metamodel, dataset = names(split_data))

```

## 4. Immune Infiltration Analysis

```{r}
conflicts_prefer(dplyr::filter)  # Resolve package conflicts

# Perform immune deconvolution
devo <- TME_deconvolution_all(split_data)

# Plot immune infiltration heatmap for a specific model
immune_heatmap = immuno_heatmap(res,
                                devo,
                                model_name = "Lasso + SuperPC",
                                dataset = "train")
png("./Mime/Ouput_27k/immune_heatmap.png", width = 18, height = 8, units = "in", res = 300)
draw(immune_heatmap)
dev.off()
```




